Hierarchical Time Series Forecasting in Emergency Medical Services

### 2022-11-18 letter/cover_letter.qmd
cover letter and title page draft

We would be grateful if you could consider for publication our manuscript entitled "Hierarchical Time Series Forecasting in Emergency Medical Services" as an Original Research article in Annals of Emergency Medicine.

Time series forecasting problems at Emergency Medical Services have often inherent hierarchical and grouped structures. For example the ambulance demand (e.g. arrival calls, verified incidents, attended incidents) in the country could be disaggregated by region, then within each region by health board, within each health board, by station/hospital, and so on down to the post code area. Demand could also be disaggrgated by other attributes that such as priority (e.g Red, Amber, Green) or nature of incidents, .

In this paper, we aim at filling various gaps identified in the literature, and our contributions to the literature are summarised as follows

>>>>

We would be grateful if you could consider for publication our manuscript entitled _Hierarchical Time Series Forecasting in Emergency Medical Services_ as an Original Research article in Annals of Emergency Medicine.

Time series forecasting problems at Emergency Medical Services have often inherent hierarchical and grouped structures. For example the ambulance demand (e.g. arrival calls, verified incidents, attended incidents) in the country could be disaggregated by region, then within each region by health board, within each health board, by station/hospital, and so on down to the post code area. Demand could also be disaggrgated by other attributes that such as priority (e.g Red, Amber, Green) or nature of incidents. Despite the popularity of such data structure in EMS, the common practice is to produce point forecast for each time series independently and ignore the structure. This practice may lead to the lack of coordination and possibly undesirable and conflicting outcomes. Practices also rely on point forecasts, which does not  acknowledge forecast uncertainty.

In this paper, we aim at filling these gaps identified in the academic literature and practice. We produce probabilistic forecasts of daily incident attended up to 84 days ahead using hierarchical forecasting methods for using state of the art forecasting methods. We apply hierarchical forecasting methods to produce a probabilistic and coherent forecast of attended incidents at multiple levels of aggregation. We use Exponential Smoothing State Space models, Poisson regression based on General Linear Models and tscount models, a Naive empirical distribution, and the ensemble model to generate base forecasts. These forecasting models collectively account for i) the spline trend, ii) multiple seasonalities, iii) holiday effects, and  iv) serial dependencies in the time series. Our study is fully reproducible and we provide data and code through a Github repository enabling reproduction and refinement of the proposed approach and benchmarks. The proposed approach could also be generalised to forecast requirements in other healthcare services, such as admissions in Emergency Departement, or call volumes in clinical desk services.

### 2022-11-18 manuscript/title_page.qmd
cover letter and title page draft

- **Author contributions:** BRT and RJH conceptualize the the study and designed the experiment. BRT obtained a funding to meet RHH at Monash University. BRT and RJH provided the analysis of the data. RJH developed the models; ML chaired the data oversight committee. BRT drafted the manuscript, and RJH contributed substantially to its revision. BRT takes responsibility for the paper as a whole.

>>>>

- **Author contributions:** BRT and RJH conceptualize the the study and designed the experiment. BRT obtained a funding to meet RHH at Monash University in Melbourne. BRT and RJH provided the analysis of the time series data. RJH developed forecasting models and hierarchical methods; BRT and RJH use R programming to conduct analysis. BRT drafted the manuscript, and RJH contributed substantially to its revision. BRT takes responsibility for the paper as a whole.
